---
title: "final project"
author: "Lizbet Ortiz"
output: 
  html_document:
    code_folding: hide
---

Introduction 

Cancer Alley in Louisiana is an 85-mile-long region with a concentration of petrochemical and industrial plants along the Mississippi River, contributing to the area's high incidence of cancer and other illnesses. The area's population of approximately 45,000 residents is predominantly low-income, and African Americans, and they have been advocating for the reduction of emissions from these facilities that pollute the air, water, and soil in their surrounding homes. 

The Louisiana Department of Environmental Quality (LDEQ) emissions Inventory data set from 1991 to 2022 provides information about the amount and type of pollutants released by industrial facilities in Louisiana. This dataset includes facility location and emissions data in pounds or tons per year. Our team analyzed this dataset to better understand the impact of emissions from these facilities on the health and well-being of the local populace, as opposed to the economic benefits touted by these businesses.



How are chemical emissions quantified?

The LDEQ employs multiple methods to quantify emissions inventory in their public datasets, including direct measurement, calculation, and modeling. 

Direct measurement: Utilizes specialized instruments to measure emissions directly from industrial and power plants.

Calculation: Involves estimating emissions from various sources using emission factors and activity data. For example,emission factors represent the amount of pollutants emitted per unit of activity, such as the amount of carbon dioxide released per unit of fossil fuel burned.

Modeling: Utilizes computer models to estimate emissions based on various inputs, such as meteorological data and land use data, that can affect emissions.


Data Limitations

The LDEQ has identified limitations in emissions reports, which can impact the accuracy of measuring pollution from companies.

Outdated equipment and incomplete data may result in underestimating or missing pollutants.
Companies may also underestimate their pollution to avoid penalties or regulations.

Furthermore, emissions reports only cover a limited range of pollutants, which can exclude other harmful substances. These limitations may have adverse public health and economic consequences. These limitations are particularly significant in Louisiana, which has a high concentration of industrial facilities emitting pollutants. 

Additionally, limitations our team found within this dataset are

It's worth noting that this data set only includes facilities with permits subject to LAC 33:III.919 and/or LAC 33:III.510. As a result, some data may be missing, which may lead to the assumption of non-contribution to a pollutant for the purpose of analysis. 
There are also gaps in the data due to some facilities not reporting every year. 

For more detailed information about these limitations, refer to the LEDQ document linked above.




Cancer Causing Chemicals

The NIH, a division of the U.S. Department of Health and Human Services, conducts vital medical research to enhance health and extend lives. The NIH recognizes various carcinogenic pollutants, including:

Benzene: A highly flammable, colorless compound utilized to manufacture synthetic fibers, plastics, and other items. It's also found in tobacco smoke and gasoline. Exposure to benzene has been associated with leukemia, lymphoma, and other blood cancers.

Butadiene: A colorless gas used in the production of synthetic rubber and other materials, and exposure to this chemical has been linked to an increased risk of leukemia and other blood cancers.

Ethylene: A gas used in the production of plastics, rubber, and other materials, and exposure to this chemical has been linked to an increased risk of breast cancer, leukemia, and lymphoma.

Formaldehyde: A colorless gas used to generate resins, textiles, and other products, also found in tobacco smoke and other sources. Leukemia and other cancers have been linked to formaldehyde exposure.

Vinyl chloride: A colorless gas used to manufacture PVC plastic products like pipes, wire coatings, and flooring. Vinyl chloride exposure has been linked to liver cancer and other cancers.


It's important to note that the risks associated with exposure to these chemicals can vary depending on the dose, duration, and other factors. Cancer is a complex illness with numerous causes, and exposure to these substances is just one of several factors that may contribute to cancer risk.



Project Focus: Ascension and Calcasieu Counties

The EIP's 2018 report highlighted that only ten counties in the United States contributed over a third (34%) of total toxicity-weighted air emissions. Intriguingly, Ascension and Calcasieu, both located in Louisiana, were among these counties. 

After analyzing the original dataset, we focused on the 15 chemical pollutants related to cancer, out of the initial 173. The pollutants were reported in Louisiana's dataset, and include 
1,3-Butadiene
Arsenic (and compounds)
Asbestos
Benzene
Benzidine
Beryllium
Cadmium (and compounds)
Chromium VI (and compounds)
Ethylene oxide
Formaldehyde
Nickel (and compounds)
Nickel (refinery dust)
Sulfuric acid
Trichloroethylene
Vinyl chloride. 

It's worth noting that the original dataset covered the period between 1991 and present. However, we focused on the 2000-2018 period to explore trends further in line with the EIP's 2018 report.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(purrr)
library(rvest)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(dplyr)

# Get data from local cvs file (data was downloaded from https://www.deq.louisiana.gov/page/eric-public-reports)
data_set_before_2015 <- read_excel("Annual_Certified_Emissions_before_2015.xlsx") 
data_set_after_2015 <- read_excel("Annual_Certified_Emissions_after_2015.xlsx")
#View(dataset)
```

```{r}
# Reusable Variables
graph_colors <- c("#999999", 
                  "#E69F00", 
                  "#56B4E9", 
                  "#009E73", 
                  "#F0E442", 
                  "#0072B2")
legend_variables_labels <- c("Others",
                             "Benzene",
                             "Butadiene",
                             "Ethylene",
                             "Formaldehyde",
                             "Vinyl")
```


The National Institutes of Health (NIH) is a part of the U.S. Department of Health and Human Services and is the nation's medical research agency that makes important discoveries that improve health and save lives.
Some of the most commonly recognized pollutants that contribute to cancer, as identified by the NIH, include:

1. Benzene: This is a colorless, highly flammable chemical used to produce plastics, synthetic fibers, and other products. It is also found in gasoline and tobacco smoke. Benzene exposure has been linked to leukemia, lymphoma, and other blood cancers.

2. Asbestos: This is a naturally occurring mineral fiber widely used in building materials such as insulation, roofing, and flooring until its health hazards were recognized. Asbestos exposure has been linked to lung cancer, mesothelioma, and other cancers.

3. Arsenic: This is a naturally occurring element found in groundwater and soil. Arsenic exposure can occur through drinking water, food, and air pollution. Arsenic exposure has been linked to skin, lung, liver, bladder, and kidney cancer.

4. Vinyl chloride: This is a colorless gas used to produce polyvinyl chloride (PVC) plastic products such as pipes, wire coatings, and flooring. Vinyl chloride exposure has been linked to liver cancer and other cancers.

5. Formaldehyde: This is a colorless gas used to produce resins, textiles, and other products and is also found in tobacco smoke and other sources. Formaldehyde exposure has been linked to leukemia and other cancers.


It's important to note that many other chemical pollutants have been linked to cancer and that the same risks associated with exposure to these chemicals can vary depending on the dose, duration, and other factors. Cancer is a complex disease with many causes, and exposure to these chemicals is just one of many factors that can contribute to cancer risk. We decided to create a new dataset that shows the 15 chemical pollutants that have been to cancer in the state of Louisiana. Another important note is that the dataset raged from 1991-Present, but for simplistic purposes we decided to focus on 2000-2018.

```{r}
# Select pollutants/variables that contribute the most to cancer
data_set_before_2015 <- data_set_before_2015 %>% 
  select(REPORTING_YEAR,
        AI_NAME, 
        PARISH_OR_COUNTY_DESC,
        `1,3-Butadiene`,
        `Arsenic (and compounds)`,
        `Asbestos`,
        `Benzene`,
        `Benzidine`,
        `Beryllium (Table 51.1)`,
        `Cadmium (and compounds)`,
        `Chromium VI (and compounds)`,
        `Ethylene oxide`,
        `Formaldehyde`,
        `Nickel (and compounds)`,
        `Nickel (refinery dust)`,
        `Sulfuric acid`,
        `Trichloroethylene`,
        `Vinyl chloride`
        )

# Select pollutants/variables that contribute the most to cancer
data_set_after_2015 <- data_set_after_2015 %>% 
  select(REPORTING_YEAR,
        AI_NAME, 
        PARISH_OR_COUNTY_DESC,
        `1,3-Butadiene`,
        `Arsenic (and compounds)`,
        `Asbestos`,
        `Benzene`,
        `Benzidine`,
        `Beryllium (Table 51.1)`,
        `Cadmium (and compounds)`,
        `Chromium VI (and compounds)`,
        `Ethylene oxide`,
        `Formaldehyde`,
        `Nickel (and compounds)`,
        `Nickel (refinery dust)`,
        `Sulfuric acid`,
        `Trichloroethylene`,
        `Vinyl chloride`
        )
```


```{r}
# Join both tables to have a single data set
data_set <- data_set_before_2015 %>% 
  full_join(data_set_after_2015, 
            by = join_by(REPORTING_YEAR,
                        AI_NAME, 
                        PARISH_OR_COUNTY_DESC,
                        `1,3-Butadiene`,
                        `Arsenic (and compounds)`,
                        `Asbestos`,
                        `Benzene`,
                        `Benzidine`,
                        `Beryllium (Table 51.1)`,
                        `Cadmium (and compounds)`,
                        `Chromium VI (and compounds)`,
                        `Ethylene oxide`,
                        `Formaldehyde`,
                        `Nickel (and compounds)`,
                        `Nickel (refinery dust)`,
                        `Sulfuric acid`,
                        `Trichloroethylene`,
                        `Vinyl chloride`)) %>% 
  arrange(AI_NAME)
```
The Environmental Integrity Project (EIP) is a nonpartisan, nonprofit watchdog organization that advocates for effective enforcement of environmental laws. Comprised of former EPA enforcement attorneys, public interest lawyers, analysts, investigators, and community organizers, EIP has three goals:

- To illustrate through objective facts and figures how the failure to enforce or implement environmental laws increases pollution and harms public health;

- To hold federal and state agencies, as well as individual corporations, accountable for failing to enforce or comply with environmental laws; and

- To help local communities obtain the protections of environmental laws.

According to the EIP, 


```{r}
# Creating `ascension` table and filter data based on `Ascension` county
ascension <- data_set %>% 
  filter(PARISH_OR_COUNTY_DESC == "Ascension") %>% 
  arrange(AI_NAME)

# Creating `calcasieu` table and filter data based on `Calcasieu` county
calcasieu <- data_set %>% 
  filter(PARISH_OR_COUNTY_DESC == "Calcasieu") %>% 
  arrange(AI_NAME)
```

```{r}
# Selecting facilities that contributed to the production of at least one pollutant
ascension2 <- ascension %>% filter(rowSums(is.na(ascension)) != ncol(ascension) - 3)

# Remove Columns by Index
ascension2 <- ascension2[-c(2:3)]


calcasieu2 <- calcasieu %>% filter(rowSums(is.na(calcasieu)) != ncol(calcasieu) - 3)
calcasieu2 <- calcasieu2[-c(2:3)]

# Replace NA values with 0
ascension2[is.na(ascension2)] <- 0
calcasieu2[is.na(calcasieu2)] <- 0
```

```{r}
# Table that holds the Ascension total number of facilities by year 
ascension_total_facilities <- ascension %>% 
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>% 
  group_by(REPORTING_YEAR) %>% 
  summarise(TOTAL_FACILITIES = n() ) 

# Table that holds the Ascension total number of facilities that contributes to at least one cancer pollutant by year
ascension_pollutant_facilities <- ascension2 %>% 
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>% 
  group_by(REPORTING_YEAR) %>% 
  summarise(POLLUTANTL_FACILITIES = n() ) 

# Table that holds Ascension pollutant and non-pollutant facilities
ascension_facilities <- ascension_total_facilities %>% 
  left_join(ascension_pollutant_facilities, 
            join_by(REPORTING_YEAR))
```


```{r}
# Plotting Ascension findings
ascension_facilities %>% mutate(NON_POLLUTANT_FACILITIES = TOTAL_FACILITIES - POLLUTANTL_FACILITIES) %>% 
  select(REPORTING_YEAR,
         NON_POLLUTANT_FACILITIES,
         POLLUTANTL_FACILITIES) %>% 
  pivot_longer(cols = 2:3,
               names_to = "type",
               values_to = "total") %>% 
  ggplot(aes(x = factor(REPORTING_YEAR), y = total, fill = type)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Number of Facilities in Ascension (2000-2018)",
       x = "Year",
       y = "Total Facilities") + 
 theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_fill_manual(name="",
                    values=c("#999999", 
                             "#E69F00"), 
                    labels = c("Non-pollutant facilities", 
                               "Pollutant Facilities"))
  
```
Table 1. Displays a bar graph that categorizes the total number of facilities in Ascension Parish according to their pollutant and non-pollutant status for the years 2000-2018 and grouped by year.

This graph provides a visual representation of Ascension Parish's facilities, distinguishing between those that produce pollutants and those that do not.

Non-pollutant facilities remained stable throughout the years, with a slight uptick in 2018.
Pollutant facilities saw a significant increase over the same period. 

This information is useful for identifying areas of concern regarding pollution control and prevention efforts and gaining an understanding of the facilities' spatial distribution.





```{r}
# Table that holds the Calcasieu total number of facilities by year 
calcasieu_total_facilities <- calcasieu %>% 
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>% 
  group_by(REPORTING_YEAR) %>% 
  summarise(TOTAL_FACILITIES = n() ) 

# Table that holds the Calcasieu total number of facilities that contributes to at least one cancer pollutant by year
calcasieu_pollutant_facilities <- calcasieu2 %>% 
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>% 
  group_by(REPORTING_YEAR) %>% 
  summarise(POLLUTANTL_FACILITIES = n() ) 

# Table that holds Calcasieu pollutant and non-pollutant facilities
calcasieu_facilities <- calcasieu_total_facilities %>% 
  left_join(ascension_pollutant_facilities, 
            join_by(REPORTING_YEAR))
```


```{r}
# Plotting Calcasieu findings
calcasieu_facilities %>% mutate(NON_POLLUTANT_FACILITIES = TOTAL_FACILITIES - POLLUTANTL_FACILITIES) %>% 
  select(REPORTING_YEAR, 
         POLLUTANTL_FACILITIES, 
         NON_POLLUTANT_FACILITIES) %>% 
  pivot_longer(cols = 2:3,
               names_to = "type",
               values_to = "total") %>% 
  ggplot(aes(x = factor(REPORTING_YEAR), y = total, fill = type)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Number of Facilities in Calcasieu (2000-2018)",
       x = "Year",
       y = "Total Facilities") + 
 theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_fill_manual(name="",
                    values=c("#999999", 
                             "#E69F00"), 
                    labels = c("Non-pollutant facilities", 
                               "Pollutant Facilities"))
  
```
Table 2. Displays a bar graph that categorizes the total number of facilities in Calcasieu County according to their pollutant and non-pollutant status for the years 2000-2018 and grouped by year.

This graph provides a visual representation of Calcasieu's facilities, distinguishing between those that produce pollutants and those that do not.

Non-pollutant facilities had a dramatic downfall in the year 2016 and remained slightly constant.
Pollutant facilities saw a slight increase over the same period. 


It's not clear that there was a significant drop in the number of chemical facilities in Calcasieu Parish, Louisiana after 2005. However, it's worth noting that the chemical industry has faced a number of challenges in recent years, including changes in market conditions and regulatory requirements, as well as the impacts of natural disasters such as experiencing significant impacts from Hurricane Rita, which made landfall in the area less than a month after Hurricane Katrina. 

Rita was a Category 3 hurricane that caused widespread damage and power outages in southwest Louisiana, including Calcasieu Parish. The storm surge and high winds caused significant damage to buildings and infrastructure, and many residents were displaced from their homes for weeks or months.

For example, the Citgo refinery in Lake Charles, which is located in Calcasieu Parish, had to shut down for several weeks after the storm due to flooding and power outages. However, many other companies in the area were able to quickly resume operations and continue producing essential products such as gasoline and other fuels to support recovery efforts in the region.

One potential factor that may have contributed to a decline in the number of facilities in the area is increasing competition from other regions, particularly those with access to cheaper feedstocks or lower operating costs. For example, the shale gas boom in Texas and other parts of the country has led to significant investments in chemical production facilities in those areas.

Overall, it's likely that a combination of factors has contributed to changes in the number and composition of chemical facilities in Calcasieu Parish and other industrial regions over time.


```{r}
#calculated the amount for each pollutant in ascension county from 2000 to 2018

total_ascension <- ascension2 %>%
  # only show years from 2000 to 2018
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>%
  group_by(REPORTING_YEAR) %>%
  summarise( total_benzene = sum(Benzene),
             total_butadiene = sum(`1,3-Butadiene`),
             total_ethylene = sum(`Ethylene oxide`),
             total_formaldehyde = sum(Formaldehyde),
             total_vinyl_chloride = sum(`Vinyl chloride`),
             others = sum(`Arsenic (and compounds)` +
                           Asbestos +
                           Benzidine + 
                           `Beryllium (Table 51.1)` + 
                           `Cadmium (and compounds)` +
                           `Chromium VI (and compounds)` + 
                           `Ethylene oxide` + 
                           Formaldehyde +  
                           `Nickel (refinery dust)` +
                           `Nickel (and compounds)` +
                           Trichloroethylene)) 

total_ascension <- total_ascension %>% 
  pivot_longer(cols = 2:7,
               names_to = "pollutant",
               values_to = "total") %>% 
  filter(`total` != 0)  %>% # Filter out rows where total is 0
  mutate(group = ifelse(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2006, 
                        1,
                        ifelse(REPORTING_YEAR >= 2007 & REPORTING_YEAR <= 2012, 
                               2,
                               3)))

# Creating subsets for different year range
total_ascension_subset1 <- filter(total_ascension, `group` == 1)
total_ascension_subset2 <- filter(total_ascension, `group` == 2)
total_ascension_subset3 <- filter(total_ascension, `group` == 3)
```

```{r}
total_ascension %>%  
  ggplot() +
  geom_bar(stat = "identity", mapping = aes(x = factor(REPORTING_YEAR), 
                                            y = total, 
                                            fill = pollutant), 
           position = "fill") +
   scale_fill_manual(name = "Names of Pollutant", 
                     values = graph_colors,
                     labels = legend_variables_labels) +
  labs(title = "Ascension Prevalence of Pollutant Concentration (2000-2018)",
       x = "Year",
       y = "Total Concentration") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) 
```
Table 3. Displays a bar graph that categorizes the six major pollutants released in the Ascension parish for the years 2000-2018.

The graph reveals a fluctuating pattern of pollutant concentrations over the years.
Benzene and Ethylene oxide with some pollutants showing a decreasing trend. 
Formaldehyde, 1,3-Butadiene, Vinyl chloride, and Others show an increasing trend 

Furthermore, the "others" category, a combination of various pollutants, exhibits consistently high concentration levels throughout the years.
In conclusion, the graph emphasizes the ongoing need for monitoring and managing pollutant concentrations in the Ascension parish to ensure that they remain within safe limits.


```{r}
# Plotting 3 Ascension subsets to visualize every pollutant by year better

# Subset 1
total_ascension_subset1 %>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, y = total, fill = pollutant), 
           position = position_dodge(0.9)) +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  labs(title = "Ascension Prevalence of Pollutant Concentration (2000-2006)",
       x = "Year",
       y = "Total Concentration (pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)

# Subset 2
total_ascension_subset2 %>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, y = total, fill = pollutant), 
           position = position_dodge(0.9)) +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  
  labs(title = "Ascension Prevalence of Pollutant Concentration (2007-2012)",
       x = "Year",
       y = "Total Concentration (pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)

# Subset 3
total_ascension_subset3 %>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, 
                         y = total, 
                         fill = pollutant), 
           position = position_dodge(0.9)) +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  
  labs(title = "Ascension Prevalence of Pollutant Concentration (2013-2018)",
       x = "Year",
       y = "Total Concentration (pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)

```





```{r}
#calculated the amount of each pollutant in Calcasieu county from 2000 to 2018
total_calcasieu <- calcasieu2 %>%
  # only show years from 2000 to 2018
  filter(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2018) %>%
  group_by(REPORTING_YEAR) %>%
  summarise( total_benzene = sum(Benzene),
             total_butadiene = sum(`1,3-Butadiene`),
             total_ethylene = sum(`Ethylene oxide`),
             total_formaldehyde = sum(Formaldehyde),
             total_vinyl_chloride = sum(`Vinyl chloride`),
             others = sum(`Arsenic (and compounds)` +
                           Asbestos +
                           Benzidine + 
                           `Beryllium (Table 51.1)` + 
                           `Cadmium (and compounds)` +
                           `Chromium VI (and compounds)` + 
                           `Ethylene oxide` + 
                           Formaldehyde +  
                           `Nickel (refinery dust)` +
                           `Nickel (and compounds)` +
                           Trichloroethylene))  
 
total_calcasieu <- total_calcasieu %>% 
  pivot_longer(cols = 2:7,
               names_to = "pollutant",
               values_to = "total") %>% 
  filter(`total` != 0)  %>% # Filter out rows where total is 0
  mutate(group = ifelse(REPORTING_YEAR >= 2000 & REPORTING_YEAR <= 2006, 
                        1,
                        ifelse(REPORTING_YEAR >= 2007 & REPORTING_YEAR <= 2012, 
                               2,
                               3)))
  # Creating subsets for different year range
total_calcasieu_subset1 <- filter(total_calcasieu, `group` == 1)
total_calcasieu_subset2 <- filter(total_calcasieu, `group` == 2)
total_calcasieu_subset3 <- filter(total_calcasieu, `group` == 3)
```



```{r}
total_calcasieu %>%  
  ggplot() +
  geom_bar(stat = "identity", 
           mapping = aes(x = factor(REPORTING_YEAR), 
                         y = total, 
                         fill = pollutant), 
           position = "fill") +
   scale_fill_manual(name = "Names of Pollutant", 
                     values = graph_colors,
                     labels = legend_variables_labels) +
  labs(title = "Calcasieu Prevalence of Pollutant Concentration (2000-2018)",
       x = "Year",
       y = "Total Concentration") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1))
```
Table 4. Displays a bar graph that categorizes the six major pollutants released in the Calcasieu parish for the years 2000-2018.

The graph reveals a fluctuating pattern of pollutant concentrations over the years.
Benzene, Butadiene and Others have the highest concentrations throughout the years.
Ethylene has the smallest concentration and appears to be decreasing throughout the years.
Formaldehyde and Vinyl have a steady concentration.

Overall, the graph provides a visual representation of the prevalence of pollutant concentrations for the specified time frame and pollutants. It can help in identifying trends and patterns of pollution levels, which can inform policy decisions and environmental regulations.


```{r}
# Plotting 3 Ascension subsets to visualize every pollutant by year better

# Substet 1 
total_calcasieu_subset1 %>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, 
                         y = total/10000, 
                         fill = pollutant), 
           position = position_dodge(0.9)) +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  labs(title = "Calcasieu Prevalence of Pollutant Concentration (2000-2006)",
       x = "Year",
       y = "Total Concentration (10,000 pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)

# Subset 2
total_calcasieu_subset2 %>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, 
                         y = total/10000, 
                         fill = pollutant), 
           position = "dodge") +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  
  labs(title = "Calcasieu Prevalence of Pollutant Concentration (2007-2012)",
       x = "Year",
       y = "Total Concentration (10,000 pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)


# Substet 3
total_calcasieu_subset3%>%  
  ggplot() +
  geom_col(mapping = aes(x = REPORTING_YEAR, 
                         y = total/10000, 
                         fill = pollutant), 
           position = "dodge") +
  scale_fill_manual(name = "Names of Pollutant", 
                    values=graph_colors,
                    labels = legend_variables_labels) +
  labs(title = "Calcasieu Prevalence of Pollutant Concentration (2013-2018)",
       x = "Year",
       y = "Total Concentration (10,000 pounds)") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1)) +
  scale_x_continuous(n.breaks = 10)

```

Conclusion

To summarize, the LDEQ emissions inventory data set offers valuable insights into the nature and extent of air pollution in Louisiana. Specifically, it provides information on the quantity and types of pollutants emitted by different sources, as well as their geographic locations and trends over time. This data can inform research, policy development, and public education efforts aimed at reducing pollution and safeguarding public health and the environment.







#IGNORE
```{random code}
#total_calcasieu %>%  
  ggplot(aes(x = factor(REPORTING_YEAR), y = total, fill = pollutant)) +
  geom_bar(stat = "identity") +
   scale_fill_manual(name = "Names of Pollutant", values= graph_colors +
  #labs(title = "Prevalence of Mean Pollutant Concentration (2000-2018)",
       x = "Year",
       y = "Mean Concentration") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


